---
title: "Airbnb"
author: "Sanjana Gupta"
date: "30/01/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
always_allow_html: yes
---

#### GitHub repo - https://github.com/sanjanagupta16/DATA598C-Data-Science-Process

Importing Libraries and setting cwd
```{r}
library(tidyverse)
library(plotly)
library(ggplot2)
library(naniar)
library(leaflet)
library(htmltools)
library(choroplethr)
library(dplyr)
library(choroplethrMaps)
library(ggplot2)
library(gridExtra)
library(ggmap)
library(htmlwidgets)
library(mapview)
library(dplyr)
library(plotly)
library(devtools)
setwd("~/Desktop/Winter 2020/Data 598C Data Science Process/Data Science Project")
```

Importing data and creating a training dataset
```{r}
data1 <- read.csv(file = "data/listings.csv", header = TRUE)
data2 <- read.csv(file = "data/calendar.csv")
data3 <- read.csv(file = "data/reviews.csv")

data_train <- data1
```

Viewing the data in different formats
```{r}
#View(data_train)

#to view data kinda nicely - uses dplyr
glimpse(data_train)
#str(data1)
#summary(data_train)
#head(data_train, 1000)
#tail(data_train, 50)
```

## Data Understanding & Data Cleanng

1. Data Cleaning - removing columns
```{r}
#cols_to_delete <- c("last_scraped", "thumbnail_url", "host_picture_url", "medium_url", "picture_url", "xl_picture_url", "host_thumbnail_url", "host_picture_url")
#select(data_train, -contains("url"))
data_train$last_scraped <- NULL
data_train$thumbnail_url <- NULL
data_train$host_picture_url <- NULL
data_train$medium_url <- NULL
data_train$picture_url <- NULL
data_train$xl_picture_url <- NULL
data_train$host_thumbnail_url <- NULL
data_train$host_picture_url <- NULL
data_train$host_url <- NULL
data_train$scrape_id <- NULL
data_train$experiences_offered <- NULL
data_train$neighborhood_overview <- NULL
data_train$host_about <- NULL
data_train$host_id <- NULL
data_train$host_verifications <- NULL
data_train$host_has_profile_pic <- NULL
data_train$host_identity_verified <- NULL
data_train$calendar_last_scraped <- NULL
data_train$license <- NULL
data_train$instant_bookable <- NULL
data_train$require_guest_profile_picture <- NULL
data_train$require_guest_phone_verification <- NULL
data_train$summary <- NULL
data_train$notes <- NULL
#Down to 69 columns from initial 92

rownames(data_train) <- NULL

#Viewing the data
#glimpse(data_train)
#summary(data_train$is_location_exact)
```

2. Converting Data type for price to numeric
```{r}
data_train$price <- sub("\\$","",data_train$price)
data_train$price <- sub(",","",data_train$price)
data_train$price <- as.integer(data_train$price)
```

3. Changing data type of categorical variables to factor
```{r}
data_train$host_response_time <- as.factor(data_train$host_response_time)
data_train$host_is_superhost <- as.factor(data_train$host_is_superhost)
data_train$neighbourhood_cleansed <- as.factor(data_train$neighbourhood_cleansed)
data_train$is_location_exact <- as.factor(data_train$is_location_exact)
data_train$property_type <- as.factor(data_train$property_type)
data_train$room_type <- as.factor(data_train$room_type)
data_train$bed_type <- as.factor(data_train$bed_type)
data_train$calendar_updated <- as.factor(data_train$calendar_updated)
data_train$cancellation_policy <- as.factor(data_train$cancellation_policy)
```

4. Changing host reponse time and extra people from character to numeric
```{r}
data_train$host_response_rate<- as.numeric(sub("%", "", data_train$host_response_rate))
data_train$host_response_rate <- data_train$host_response_rate/100
data_train$extra_people <- as.numeric(sub("\\$","",data_train$extra_people))
```

5. Checking if Price = 0 for any listings
```{r}
0 %in% data_train$price
```
All Airbnb listings have a price associated with it

6. Removing rows with NA
```{r}
#is.null(data_train$zipcode)
data_train %>% filter_all(all_vars(!is.na(.)))
data_train %>% filter_all(all_vars(complete.cases(.)))  
```

7. Removing rows with null values
```{r}
data_train %>% filter_all(all_vars(!is.null(.)))
data_train %>% filter_all(all_vars(complete.cases(.)))  
```

8. Getting a summary of all fields
```{r}
summary_df <- data.frame(summary(data_train))
#write.csv(summary_df, "summary.csv")
```

9. Checking for missing values and plotting the data
```{r}
check_missing<- function(x){
  if (is.character(x)) sum(x=="") else sum(is.na(x))
}
NMISS<-data.frame(nmiss=sapply(data_train,check_missing))
#write.csv(NMISS, "missing.csv")

#plotting missing data
missingdata <- data_train
missingdata[missingdata == ""] <- NA
missingdata <- missingdata %>% select(host_is_superhost, review_scores_rating,
                                      host_response_time, name, host_since, zipcode, latitude,longitude,
                                      host_location, host_response_rate,neighbourhood_cleansed,
                                      property_type, price,weekly_price, monthly_price, bedrooms, bathrooms)
gg_miss_var(missingdata) + labs(title = "Missing Values Plot",y = "Missing Data", x = "Attribute")  + theme_bw() + theme(text=element_text(color = "dark blue")) 
```

10. Replacing missing values
```{r}
data_train$review_scores_rating <- ifelse(is.na(data_train$review_scores_rating)==T,97,data_train$review_scores_rating)
data_train$host_response_rate <- ifelse(is.na(data_train$host_response_rate)==T,97,data_train$host_response_rate)

#Using same plot as before to check the values
gg_miss_var(missingdata) + labs(title = "Missing Values Plot",y = "Missing Data", x = "Attribute")  + theme_bw() + theme(text=element_text(color = "dark blue")) 
```

## Data Understanding - plots

1. Analysing neighborhood grouping
```{r}
#In neighborhood group cleansed, there are 794 rows with "Other Neighborhoods"
#Need to fix this - maybe look at neighorhood cleansed column instead?
#Looking at categorical data now
sort(table(data_train$host_neighbourhood), decreasing = TRUE)
head(subset(data_train, select = 'neighbourhood_group_cleansed'))
factor(data_train$neighbourhood_group_cleansed)
count(data_train, 'neighbourhood_group_cleansed')
count(data_train, 'neighbourhood_cleansed')

#Getting a count of listings per neighborhood
#write.csv(neigborhood_data, "neighborhoodcount.csv")
#count(data_train$neighbourhood_cleansed)
plot(data_train$neighbourhood_cleansed, xlab = 'Neighborhoods', ylab = 'Count')
neighborhood <- (data_train$neighbourhood_cleansed)
#x <- sort(count((neighborhood)))
#arrange(data_train, neighbourhood_cleansed)
neighborhood_data <- data_train[order(data_train[,21] ),]
summary(neighborhood_data$neighbourhood_cleansed)
```

```{r}
#Trying to fix "Other Neighborhoods" value in neighborhood group cleaned column
#for(i in data_train)
 # if(i[neighbourhood_group_cleansed] == "Other neighborhoods")
  #  i[neighbourhood_group_cleansed] <-  i[neighborhood_cleansed]

#Plotting count of listings per neighborhood
neigborhood_data <- table(data_train$neighbourhood_cleansed)

barplot(sort(table(data_train$neighbourhood_group_cleansed), decreasing = TRUE), ylab = "Count",horiz = FALSE , main = "Distribution of Listings Per Neighborhood", beside = FALSE)
```

```{r}
#barplot(table(data_train$neighbourhood_group_cleansed), density = 20, ylab = "Count",horiz = FALSE , main = "Distribution of Listings Per Neighborhood", beside = FALSE) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
ggplot(data_train, aes(neighbourhood_group_cleansed)) + geom_bar(na.rm = TRUE, stat = "count") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

2. Looking at some more plots - Count of Host Listings
```{r}
#barplot(data_train$host_listings_count, main = "Count of Host Listings", xlab = "", ylab = "Count") 
ggplot(data_train, aes(host_listings_count)) + geom_bar(na.rm = TRUE) + xlim(0,21) 
```
# Most hosts have a single Airbnb listing.

3. Looking at Types of Airbnb properties
```{r}
#barchart(data_train$property_type, main = "Types of Airbnb Properties", xlab = "Count")
prop_type <- table(data_train$property_type)
barplot(table(data_train$property_type), ylab = "Count",horiz = FALSE, main = "Types of Airbnb Properties", beside = FALSE, las = 1)

ggplot(data_train, aes(property_type)) + geom_bar(na.rm = TRUE, stat = "count") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
#plot(data_train$host_is_superhost)
```

4. Prices per nNeighborhood
```{r}
table(data_train$room_type,data_train$neighbourhood_group)
```
```{r}
data_train %>% boxplot(price ~ neighbourhood_group_cleansed,data = ., main="Box Plot of price vs neighbourhood", 
                     ylab="neighbourhood", xlab="Price",horizontal=TRUE)
```

5. Number of listings per neighborhood
```{r}
ggplot(data_train, aes(x = fct_infreq(neighbourhood_cleansed), fill = room_type)) +
    geom_bar() +
    labs(title = "No. of listings by neighborhood",
         x = "Neighborhood", y = "No. of listings") +
    theme(legend.position = "bottom") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
Looks messy, want to consider looking at only the top 10 neighborhoods

6. Top 10 neighborhoods - doesn't work - need to fix this
```{r}
#top_10 <- data_train %>% group_by(neighborhood_cleansed) %>% tally()
#top_10 <- data_train %>% top_n(n = 10, wt = neighbourhood_group_cleansed) %>% arrange(neighbourhood_group_cleansed) %>% #summarize(neighbourhood_group_cleansed = n())

top_10 <- data_train[data_train$neighbourhood_group_cleansed %in% names(sort(table(data_train$neighbourhood_group_cleansed), decreasing = TRUE)[0:1]),]
#top_10 <- data_train %>% top_n(10) %>% group_by(neighbourhood_group_cleansed) %>% tally() 
table(top_10$neighbourhood_group_cleansed)
```

7. Creating a variable to find the number of days since the listing has been on Airbnb
```{r}
data_train$host_since <- as.Date(data_train$host_since)
data_train$host_since <- as.Date("2017-05-10")-data_train$host_since
```

8. Looking at outliers for price
```{r}
ggplot(data=data_train, aes(price)) + 
  geom_histogram(fill="red") + 
  labs(title="Histogram of Price") +
  labs(x="Price", y="Count")
# Percentile of price
quantile(data_train$price, c(.9, .95, .97, 0.975, 0.98, 0.99, 0.995, 0.999, 0.9999))
```
9. Room Type Analysis
# For a private room
```{r}
Private_rooms <- data_train %>% filter(room_type == "Private room")
Private_rooms$price %>% summary()
```
# For an entire apartment
```{r}
entire_house <- data_train %>% filter(room_type == "Entire home/apt")
entire_house$price %>% summary()
```
#Shared room
```{r}
shared_room <- data_train %>% filter(room_type == "Shared room")
shared_room$price %>% summary()
```

10. Average prices for each room type
```{r}
listingdf <- read.csv('data/listings.csv')
zipReviews <- listingdf %>% group_by(zipcode = zipcode) %>% summarise(avg_loc_review = mean(review_scores_location, na.rm = TRUE))

average_price <- data_train %>% group_by(room_type) %>% summarise(price = mean(price))
colnames(average_price)[1] <- "Room Type"
colnames(average_price)[2] <- "Average Price"
print((average_price))
```

11. Top 50 most expensive listings
```{r patial Data 1.3, echo=FALSE, message=FALSE, warning=FALSE}
top_df <- data_train %>% top_n(n = 50, wt = price)

# get background map
top_height <- max(top_df$latitude) - min(top_df$latitude)
top_width <- max(top_df$longitude) - min(top_df$longitude)
top_borders <- c(bottom  = min(top_df$latitude)  - 0.1 * top_height,
                 top     = max(top_df$latitude)  + 0.1 * top_height,
                 left    = min(top_df$longitude) - 0.1 * top_width,
                 right   = max(top_df$longitude) + 0.1 * top_width)

top_map <- get_stamenmap(top_borders, zoom = 12, maptype = "toner-lite")

# map of top 50 most expensive
ggmap(top_map) +
    geom_point(data = top_df, mapping = aes(x = longitude, y = latitude,
                                        col = price)) +
    scale_color_gradient(low = "blue", high = "red")
```

## Model Selection

1. Clustering
Used Clustering to compare which areas of Seattle have a higher concentration of airbnb listings.

Plotting the listings on the map of Seattle
```{r}
#Reading Listings Data
listingdf <- read.csv('data/listings.csv')
#Creating a link for each pop up

#Creating Listings across Seattle
map <- leaflet(listingdf) %>%
  addTiles() %>%
  addMarkers(~longitude, ~latitude,labelOptions = labelOptions(noHide = T),
             clusterOptions = markerClusterOptions(),
             popup = paste0("<b> Name: </b>", listingdf$name , "<br/><b> Host Name: </b>", 
                            listingdf$host_name, "<br> <b> Price/night: </b>", listingdf$price, "<br/><b> Room Type: </b>", 
                            listingdf$room_type, "<br/><b> Property Type: </b>", listingdf$property_type
             )) %>% 
  setView(-122.335167, 47.608013, zoom = 11) %>%
  addProviderTiles("CartoDB.Positron")
map
mapshot(map, url = paste0(getwd(),"/map.html"))
```
This shows listings in a clustered fashion. Clicking on a cluster zooms into the listings belonging to it.
Clicking on a pin gives you a pop up with some details about the property.
We can see that most of the listings are clustered around Downtown and Cap Hill.

2. Linear Regression
Which variables influence price?
Variables selected - bathrooms, price, property type, accomodates, bedrooms, room type
```{r}
LM_res <- summary(lm(formula = price ~ room_type + accommodates + bedrooms + bathrooms + property_type, data = data_train))
plot(price ~room_type + accommodates + bedrooms + bathrooms + property_type + factor(neighbourhood_cleansed), data = data_train) + abline(lm(formula = price ~ room_type + accommodates + bedrooms + bathrooms, data = data_train))
LM_res

```
Private and Shared rooms, accommodates, bedrooms, bathrooms have a singificant impact on price.
So does Property Type = Boat - which is weird.

Pairwise comparison
```{r}
# price vs bathrooms
p1 <- ggplot(data_train, aes(x=bathrooms, y=price)) +
  geom_point(colour = "orange", size = 1.5) +
  geom_smooth(method='lm', color='red') +
  ylab("price")

# price vs accommodates
p2 <- ggplot(data_train, aes(x=accommodates, y=price)) +
  geom_point(colour = "orange", size = 1.5) +
  geom_smooth(method='lm', color='red') +
  ylab("price")

# price vs bedrooms
p3 <- ggplot(data_train, aes(x=bedrooms, y=price)) +
  geom_point(colour = "orange", size = 1.5) +
  geom_smooth(method='lm', color='red') +
  ylab("price")

# price vs room_type
p4 <- ggplot(data_train, aes(x=room_type, y=price)) +
  geom_point(colour = "orange", size = 1.5) +
  geom_smooth(method='lm', color='red') +
  ylab("price")

# combining all plots
grid.arrange(p1, p2, p3, p4,ncol = 2)
```

3. Predicting Price using Linear Regression Model
```{r}
set.seed(100)
trainingRowIndex <- sample(1:nrow(data_train), 0.8 *nrow(data_train))
trainingData <- data_train[trainingRowIndex,]
testData <- data_train[-trainingRowIndex,]
lm_model <- lm(formula = price ~ room_type + accommodates + bedrooms + bathrooms, data = trainingData)
price_pred <- predict(lm_model, testData)

summary(lm_model)

actual_pred <- data.frame(cbind(actuals = testData$price, predicteds = price_pred))
correlation_accuracy <- cor(actual_pred)

head(actual_pred)
```

Model Fit for Linear Regression -- R^2, Residual Analysis, F-statistic

4. ANOVA - to compare categorical data
```{r}
AOV_res <- summary(aov(price ~ factor(property_type) + factor(neighbourhood_cleansed), data = data_train))
AOV_res
```
As we can see, property type and neighborhood do have a significant impact on price since the p-value is much smaller than the significance level.
